Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-21 paper.tex
save

The author disambiguation solution proposed in this work, along with its
enhancements, are evaluated on data extracted from the INSPIRE portal
\citep{gentil2009information}, a digital library for scientific literature in
high-energy physics. Overall, the portal holds more than 1 million
publications, forming in total a set ${\cal S}$ of more than 10 millions
signatures. Out of these, around 13\% have been \textit{claimed} by their
original authors or marked as such by professional curators. Together, they
constitute a trusted set $\bar{\cal C}$ of 15388 distinct individuals sharing
36340 unique author names. This data covers  several decades in time and dozens of
author nationalities worldwide.

Following INSPIRE terms of use, the data is released
online\footnote{\url{https://github.com/glouppe/paper-author-disambiguation}}
under the CC0 license. To the best of our knowledge, data of this size and
coverage is the first to be publicly released in the scope of author
disambiguation research.

% source of the data, labels in the data, etc
% release the data

@@

\glnote{By Gilles}

@@

\glnote{To review}

The starting point of our analysis is an algorithm that blocks the signatures using the last
name and first initial strategy, samples the pairs in balanced way, uses gradient boosting
with all features included for learning the similarity function and clusters using
semi-supervised hierarchical approach.

\begin{table}[H]
\caption{Base experiment evaluation)}
\centering
\begin{tabular}{|r|r|l|l|l|}
  \hline
  Evaluation & Set & Precision & Recall & F-score \\
  \hline
  Pairwise & test & 0.9948 & 0.9738 & 0.9842 \\
  \hline
  Pairwise & train & 0.9976 & 0.9747 & 0.986 \\
  \hline
  Pairwise & overall & 0.9952 & 0.9739 & 0.9844 \\
  \hline
  $B^3$ & test & 0.9901 & 0.976 & 0.983 \\
  \hline
  $B^3$ & train & 0.997 & 0.979 & 0.9879 \\
  \hline
  $B^3$ & overall & 0.9907 & 0.9761 & 0.9834 \\
  \hline
\end{tabular}
\end{table}

@@

\glnote{To review}

Let us introduce a term \textit{blocking limited maximum recall} to describe the highest
recall that the algorithm can theoretically reach after blocking is done.
Formally, for the set of signatures ${\cal S}$, blocking function $F: \mathcal{S}
\rightarrow \mathcal{B}$, where $ \mathcal{B} = \{s_{0}, \ldots{} , s_{N-1}\}$ is a
partition of $\mathcal{S} $ , the recall function $\mathcal{R}: \mathcal{C}_{gold},
\mathcal{C}_{predicted} \rightarrow \langle0,1\rangle$ where $\mathcal{C}_{gold}$ is the
partition of $\mathcal{S}$ where for every author there is a set in $\mathcal{C}_{gold}$
that contains only all the signatures of this author, $\mathcal{C}_{predicted}$ is the
partition of $\mathcal{S}$ output by the algorithm, and $\mathcal{Q} = \{\cup_{i=0}^{N-1}
q_i $ : $\forall_{i \in \langle0,N-1\rangle} q_i$ is a partition of $s_i$\}. Then
\textit{blocking limited maximum recall} is defined as $\max_{q \in \mathcal{Q}}
\mathcal{R}(\mathcal{C}_{gold}, q)$. For sanely defined recall functions, such as $B^{3}$
recall or \textit{pairwise} recall \msnote{Further explanation needed, why it is
equivalent?}, it is equivalent to $\mathcal{R}(\mathcal{C}_{gold}, \mathcal{B})$.

The \textit{blocking limited maximum recall} for the last name and first initial blocking
and $B^3$ scoring equals 0.9827. Comparing it to the $B^3$ recall of the base experiment,
one can see that a huge part of errors comes from the blocking step which forbids some of
the pairs to be merged together. To prevent that, we tried the phonetic blocking algorithm
using three different phonetic algorithms: Double Metaphone \citep{doublemetaphone}, NYSIIS
\citep{nysiis} and Soundex \citep{Soundex}. As Double Metaphone can return two results for a
string, we always discarded the second one when we used this algorithm.

\begin{table}[H]
\caption{Blocking limited maximum recalls (BLMR) on the whole dataset without splitting over
the first given name initial}
\centering
\begin{tabular}{|r|l|l|}
  \hline
  blocking function & BLMR with $B^{3}$ recall & BLMR with pairwise recall \\
  \hline
  last name and first initial & 0.9827 & 0.9776 \\
  \hline
  phonetic using double metaphone & 0.9967 & 0.9953 \\
  \hline
  phonetic using nysiis & 0.9961 & 0.9951 \\
  \hline
  phonetic using soundex & 0.9966 & 0.9953 \\
  \hline
\end{tabular}
\end{table}

\msnote{Mention two new features for this experiments, other than that everything is the
same as in the base experiment}

\begin{table}[H]
\caption{Phonetic algorithm performance}
\centering
\begin{tabular}{|r|r|r|l|l|l|}
  \hline
  Phonetic Algorithm & Split over initial & Evaluation & Precision & Recall & F-score \\
  \hline
  Double Metaphone & yes & pairwise & 0.9934 & 0.982 & 0.9876 \\
  \hline
  Double Metaphone & yes & $B^3$ & 0.9867 & 0.983 & 0.9848 \\
  \hline
  Double Metaphone & no & pairwise & - & - & - \\
  \hline
  Double Metaphone & no & $B^3$ & - & - & - \\
  \hline
  NYSIIS & yes & pairwise & 0.9942 & 0.9818 & 0.9879 \\
  \hline
  NYSIIS & yes & $B^3$ & 0.9884 & 0.9828 & 0.9856 \\
  \hline
  NYSIIS & no & pairwise & - & - & - \\
  \hline
  NYSIIS & no & $B^3$ & - & - & - \\
  \hline
  Soundex & yes & pairwise & - & - & - \\
  \hline
  Soundex & yes & $B^3$ & - & - & - \\
  \hline
  Soundex & no & pairwise & - & - & - \\
  \hline
  Soundex & no & $B^3$ & - & - & - \\
  \hline
\end{tabular}
\end{table}

\msnote{Fill the table}

>>>>

The author disambiguation solution proposed in this work, along with its
enhancements, are evaluated on data extracted from the INSPIRE portal
\citep{gentil2009information}, a digital library for scientific literature in
high-energy physics. Overall, the portal holds more than 1 million
publications, forming in total a set ${\cal S}$ of more than 10 millions
signatures. Out of these, around 13\% have been \textit{claimed} by their
original authors or marked as such by professional curators. Together, they
constitute a trusted set $\bar{\cal C}$ of 15388 distinct individuals sharing
36340 unique author names spread within 1201763 signatures on 360066
publications. This data covers several decades in time and dozens of author
nationalities worldwide.

Following INSPIRE terms of use, the dataset $\bar{\cal C}$ is released
online\footnote{\url{https://github.com/glouppe/paper-author-disambiguation}}
under the CC0 license. To the best of our knowledge, data of this size and
coverage is the first to be publicly released in the scope of author
disambiguation research.

@@

@@

% \glnote{To review}

% The starting point of our analysis is an algorithm that blocks the signatures using the last
% name and first initial strategy, samples the pairs in balanced way, uses gradient boosting
% with all features included for learning the similarity function and clusters using
% semi-supervised hierarchical approach.

% \begin{table}[H]
% \caption{Base experiment evaluation)}
% \centering
% \begin{tabular}{|r|r|l|l|l|}
%   \hline
%   Evaluation & Set & Precision & Recall & F-score \\
%   \hline
%   Pairwise & test & 0.9948 & 0.9738 & 0.9842 \\
%   \hline
%   Pairwise & train & 0.9976 & 0.9747 & 0.986 \\
%   \hline
%   Pairwise & overall & 0.9952 & 0.9739 & 0.9844 \\
%   \hline
%   $B^3$ & test & 0.9901 & 0.976 & 0.983 \\
%   \hline
%   $B^3$ & train & 0.997 & 0.979 & 0.9879 \\
%   \hline
%   $B^3$ & overall & 0.9907 & 0.9761 & 0.9834 \\
%   \hline
% \end{tabular}
% \end{table}

@@

% \glnote{To review}

% Let us introduce a term \textit{blocking limited maximum recall} to describe the highest
% recall that the algorithm can theoretically reach after blocking is done.
% Formally, for the set of signatures ${\cal S}$, blocking function $F: \mathcal{S}
% \rightarrow \mathcal{B}$, where $ \mathcal{B} = \{s_{0}, \ldots{} , s_{N-1}\}$ is a
% partition of $\mathcal{S} $ , the recall function $\mathcal{R}: \mathcal{C}_{gold},
% \mathcal{C}_{predicted} \rightarrow \langle0,1\rangle$ where $\mathcal{C}_{gold}$ is the
% partition of $\mathcal{S}$ where for every author there is a set in $\mathcal{C}_{gold}$
% that contains only all the signatures of this author, $\mathcal{C}_{predicted}$ is the
% partition of $\mathcal{S}$ output by the algorithm, and $\mathcal{Q} = \{\cup_{i=0}^{N-1}
% q_i $ : $\forall_{i \in \langle0,N-1\rangle} q_i$ is a partition of $s_i$\}. Then
% \textit{blocking limited maximum recall} is defined as $\max_{q \in \mathcal{Q}}
% \mathcal{R}(\mathcal{C}_{gold}, q)$. For sanely defined recall functions, such as $B^{3}$
% recall or \textit{pairwise} recall \msnote{Further explanation needed, why it is
% equivalent?}, it is equivalent to $\mathcal{R}(\mathcal{C}_{gold}, \mathcal{B})$.

% The \textit{blocking limited maximum recall} for the last name and first initial blocking
% and $B^3$ scoring equals 0.9827. Comparing it to the $B^3$ recall of the base experiment,
% one can see that a huge part of errors comes from the blocking step which forbids some of
% the pairs to be merged together. To prevent that, we tried the phonetic blocking algorithm
% using three different phonetic algorithms: Double Metaphone \citep{doublemetaphone}, NYSIIS
% \citep{nysiis} and Soundex \citep{Soundex}. As Double Metaphone can return two results for a
% string, we always discarded the second one when we used this algorithm.

% \begin{table}[H]
% \caption{Blocking limited maximum recalls (BLMR) on the whole dataset without splitting over
% the first given name initial}
% \centering
% \begin{tabular}{|r|l|l|}
%   \hline
%   blocking function & BLMR with $B^{3}$ recall & BLMR with pairwise recall \\
%   \hline
%   last name and first initial & 0.9827 & 0.9776 \\
%   \hline
%   phonetic using double metaphone & 0.9967 & 0.9953 \\
%   \hline
%   phonetic using nysiis & 0.9961 & 0.9951 \\
%   \hline
%   phonetic using soundex & 0.9966 & 0.9953 \\
%   \hline
% \end{tabular}
% \end{table}

% \msnote{Mention two new features for this experiments, other than that everything is the
% same as in the base experiment}

% \begin{table}[H]
% \caption{Phonetic algorithm performance}
% \centering
% \begin{tabular}{|r|r|r|l|l|l|}
%   \hline
%   Phonetic Algorithm & Split over initial & Evaluation & Precision & Recall & F-score \\
%   \hline
%   Double Metaphone & yes & pairwise & 0.9934 & 0.982 & 0.9876 \\
%   \hline
%   Double Metaphone & yes & $B^3$ & 0.9867 & 0.983 & 0.9848 \\
%   \hline
%   Double Metaphone & no & pairwise & - & - & - \\
%   \hline
%   Double Metaphone & no & $B^3$ & - & - & - \\
%   \hline
%   NYSIIS & yes & pairwise & 0.9942 & 0.9818 & 0.9879 \\
%   \hline
%   NYSIIS & yes & $B^3$ & 0.9884 & 0.9828 & 0.9856 \\
%   \hline
%   NYSIIS & no & pairwise & - & - & - \\
%   \hline
%   NYSIIS & no & $B^3$ & - & - & - \\
%   \hline
%   Soundex & yes & pairwise & - & - & - \\
%   \hline
%   Soundex & yes & $B^3$ & - & - & - \\
%   \hline
%   Soundex & no & pairwise & - & - & - \\
%   \hline
%   Soundex & no & $B^3$ & - & - & - \\
%   \hline
% \end{tabular}
% \end{table}

% \msnote{Fill the table}
